Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome

### 2023-05-31 paper/paper.qmd
Switch 'attrib' to 'attributable' everywhere

load(here("results", "stats.RData"))
table_font_size <- 11
```

@@

```{r sample_counts}
#| label: tbl-counts
#| tbl-pos: h
#| tbl-cap: "Dataset sample counts"
#| tbl-subcap: 
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 2
format_table <- function(tab) {
  tab %>%
    kable(col.names = c(' ', 'IDSA', 'Attrib', 'All-cause', "Pragmatic"),
          format.args = list(big.mark = ',')) %>% 
    kable_styling(font_size = table_font_size)
}
read_csv(here('results','count_table_full.csv')) %>% 
  format_table()
read_csv(here('results','count_table_int.csv')) %>% 
  format_table()
```

@@

::: {#fig-flowchart fig-cap="**CDI severity definitions.**
**A)** Decision flow chart to define CDI cases as severe according to the Infectious
Diseases Society of America (IDSA) based on lab values,
the occurrence of an adverse outcome due to any cause (All-cause),
and the occurrence of disease-related complications confirmed as attributable to
CDI with chart review (Attrib).
**B)** The proportion of severe CDI cases labelled according to each definition.
An additional 'Pragmatic' severity definition uses the Attributable definition
when possible, and falls back to the All-cause definition when chart review is
not available.
<!-- TODO table (supplementary?) showing counts & frequency of positives-->
"}
![](figures/flowchart_sankey.png)
:::

::: {#fig-performance fig-cap="**Performance of ML models.**
In the left panels, models were trained on the full dataset, with different numbers of samples
available for each severity definition.
In the right panels, models were trained on the intersection of samples with all labels
available for each definition. 
Note that the intersection dataset for Attributable and Pragmatic severity have
exactly the same labels, thus identical values are expected.
**A)** 
Area under the receiver-operator characteristic curve (AUROC) for the test sets
and cross-validation folds of the training sets,
and the area under the balanced precision-recall curve (AUBPRC) for the test sets.
Nearly all pairs of definitions have significantly different performances on the
test set (P < 0.05) except for AUROC and AUBPRC of Attributable vs. Pragmatic on
the intersection dataset (as they are identical), AUROC of Attributable vs.
All-cause on the full dataset, and AUROC of Attributable vs. IDSA on the full
dataset.
**B)** Receiver-operator characteristic curves for the test sets.
Mean specificity is reported at each sensitivity value.
**C)** Balanced precision-recall curves for the test sets.
Mean balanced precision is reported at each recall value.
"}
![](figures/ml-performance.png)
:::

>>>>

load(here("results", "stats.RData"))
table_font_size <- 10
```

@@

```{r sample_counts}
#| label: tbl-counts
#| tbl-pos: h
#| tbl-cap: "Dataset sample counts"
#| tbl-subcap: 
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 2
format_table <- function(tab) {
  tab %>%
    kable(col.names = c(' ', 'IDSA', 'Attributable', 'All-cause', "Pragmatic"),
          format.args = list(big.mark = ',')) %>% 
    kable_styling(font_size = table_font_size)
}
read_csv(here('results','count_table_full.csv')) %>% 
  format_table()
read_csv(here('results','count_table_int.csv')) %>% 
  format_table()
```

@@

::: {#fig-flowchart fig-cap="**CDI severity definitions.**
**A)** Decision flow chart to define CDI cases as severe according to the
Infectious Diseases Society of America (IDSA) based on lab values,
the occurrence of an adverse outcome due to any cause (All-cause),
and the occurrence of disease-related complications confirmed as attributable to
CDI with chart review (Attributable).
**B)** The proportion of severe CDI cases labelled according to each definition.
An additional 'Pragmatic' severity definition uses the Attributable definition
when possible, and falls back to the All-cause definition when chart review is
not available.
<!-- TODO table (supplementary?) showing counts & frequency of positives-->
"}
![](figures/flowchart_sankey.png)
:::

::: {#fig-performance fig-cap="**Performance of ML models.**
In the left panels, models were trained on the full dataset, with different
numbers of samples available for each severity definition.
In the right panels, models were trained on the intersection of samples with all
labels available for each definition.
Note that the intersection dataset for Attributable and Pragmatic severity have
exactly the same labels, thus identical values are expected.
**A)** 
Area under the receiver-operator characteristic curve (AUROC) for the test sets
and cross-validation folds of the training sets,
and the area under the balanced precision-recall curve (AUBPRC) for the test sets.
Nearly all pairs of definitions have significantly different performances on the
test set (P < 0.05) except for AUROC and AUBPRC of Attributable vs. Pragmatic on
the intersection dataset (as they are identical), AUROC of Attributable vs.
All-cause on the full dataset, and AUROC of Attributable vs. IDSA on the full
dataset.
Each point is the median with tails as the 95% CI.
**B)** Receiver-operator characteristic curves for the test sets.
Mean specificity is reported at each sensitivity value, 
with ribbons as the 95% CI.
**C)** Balanced precision-recall curves for the test sets.
Mean balanced precision is reported at each recall value, 
with ribbons as the 95% CI.
"}
![](figures/ml-performance.png)
:::
